Solution-aware vs global ReLU selection: \\ partial MILP strikes back for DNN verification

### 2025-04-24 ATVA 2025/experiments.tex
final?

{\sf SAS} (even {\sf GS}$=s_{FSB}$) is much more accurate than full MILP for every reasonable time-outs considered, with $>10$ times fastest runtime at better accuracy. The reason is that even the advanced MILP solver Gurobi struggles to sort out all ReLUs. It is thus particulary important to have accurate scoring functions as {\sf SAS}, in order to optimize both accuracy and runtime. Huang \cite{DivideAndSlide} is limited in ReLUs in the previous layer, reason why it is stuck at relatively poor accuracy. The accuracy/runtime curve from {\sf GS} is closer to {\sf SAS} than the number of nodes / accuracy curve of Fig. \ref{fig_table3}. This is because some ReLU nodes deemed important by {\sf GS} are not important for accuracy, but they are also not penalizing runtime too much. Still, {\sf SAS} is faster than {\sf GS} at  every accuracy. Importantly, {\sf SAS} is more deterministic in its runtime given a number of ReLU nodes, with half the variance in runtime  over different output neurons compared with {\sf GS} for similar accuracy
(not displayed), which helps setting a number of ReLU nodes to open.

>>>>

{\sf SAS} (even {\sf GS}$=s_{FSB}$) is much more accurate than full MILP for every reasonable time-outs considered, with $>10$ times fastest runtime at better accuracy. The reason is that even the advanced MILP solver Gurobi struggles to sort out all ReLUs. It is thus particulary important to have accurate scoring functions as {\sf SAS}, in order to optimize both accuracy and runtime. Huang \cite{DivideAndSlide} is limited in ReLUs in the previous layer, reason why it is stuck at relatively poor accuracy. The accuracy/runtime curve from {\sf GS} is closer to {\sf SAS} than the number of nodes / accuracy curve of Fig. \ref{fig_table3}. This is because many ReLU nodes deemed important by {\sf GS} are not relevant for accuracy, but they are also not penalizing runtime. Still, {\sf SAS} is faster than {\sf GS} at  every accuracy. Importantly, {\sf SAS} is more deterministic in its runtime given a number of ReLU nodes, with half the variance in runtime over different output neurons compared with {\sf GS} for similar accuracy (see table \ref{SAS_GS_similarTO} in Appendix), which helps setting a number of ReLU nodes to open.

%	\begin{tabular}{|c|c|c|c|c|}
%	\hline
%	NodeID & $\UB$ by SAS& Time cost & $\UB$ by GS  & Time cost \\
%	\hline
%	0 & -1.4684063 & 66.9 & -1.3804709 & 71.6 \\
%	\hline
%	2 & -4.5550154 & 54.8 & -4.5424599 & 84.9 \\
%	\hline
%	3 & -4.4795618 & 69.9 & -4.4459532 & 71.2 \\
%	\hline
%	4 & -3.4594076 & 88.8 & -3.4784499 & 113 \\
%	\hline
%	5 & -5.5981178 & 94.8 & -5.6955818 & 93.7 \\
%	\hline
%	6 & -3.0597499 & 63.3 & -2.9797569 & 68.5 \\
%	\hline
%	7 & -4.1742691 & 73.4 & -4.0842553 & 88.6 \\
%	\hline
%	8 & 0.63875415 & 56.7 & 0.72631579 & 46.5 \\
%	\hline
%	9 & -2.3270878 & 64.6 & -2.2841141 & 69.4 \\
%	\hline
%	Average & -3.1647624
%	& 70.3
%	& -3.129414017 & 78.6
%	\\
%	\hline
%	Variance & 
%	& 184
%	&  & 359
%	\\
%	\hline
%\end{tabular}

### 2025-04-24 ATVA 2025/formula.tex
an advise sample
final?

{\color{red}
\noindent {\em Comparison:} First, the original \cite{DivideAndSlide} does not propose a formula for node $a$ two layers before $z$. 
Consider again the running example of Fig. \ref{img:FSB_example}. 
In the case $c\cdot d<0$, we have $\Utility\_\max\nolimits^z(a)=\Improve_{max}^z(b)=s_{FSB}(a)=s_{SR}(a)=0$, as $\Delta(\hat{a})=0$. In the case $c>0,d>0$, 
$\Delta(\hat{a}) = \ReLU(\sol(a))-\sol(\hat{a})$ is precise,
whereas the corresponding $\Delta_{FSB}(\hat{a}) = \LB(a) r(a)$ is only an upperbound.
The last case $c<0,d<0$ is the most extreme: 
$\Delta(\hat{b})$ adapts to the case (b),(c) in Fig. \ref{node:b} leveraging the value $\sol(b)$, which yields very different values, whereas the corresponding $\Delta_{FSB}(\hat{b})$
is always $r(b) \Delta_{FSB}(b)$. For $sol(b)<<0$, we will have 
$\Delta(\hat{b})=0<\Delta_{FSB}(\hat{b})$.
For $sol(b)>>0$, we will have 
$\Delta(\hat{b})=\Delta(b) > \frac{1}{2}\Delta_{FSB}(\hat{b})$ 
as $r(b)=\frac{1}{2}$. We can show that $\Utility$ is a safe overapproximation of 
$\Improve\_\max^z(a)$, which does not hold for $s_{FSB},s_{SR}$ (because of the case $sol(b)>>0$):
}

>>>>

\noindent {\em Comparison:} First, the original \cite{DivideAndSlide} does not propose a formula for node $a$ two layers before $z$. So we will compare {\sf SAS} with {\sf GS}.
Consider again the running example of Fig. \ref{img:FSB_example}.

In the case $c\cdot d<0$, we have $\Utility\_\max\nolimits^z(a)=\Improve_{max}^z(b)=s_{FSB}(a)=s_{SR}(a)=0$, as $\Delta(\hat{a})=0$.

In the case $c>0,d>0$, 
$\Delta(\hat{a}) = \ReLU(\sol(a))-\sol(\hat{a})$ is precise,
whereas the corresponding $\Delta_{FSB}(\hat{a}) = \LB(a) r(a)$ is only an upperbound.

The last case $c<0,d<0$ is the most extreme: 
$\Delta(\hat{b})$ adapts to the case (b),(c) in Fig. \ref{node:b} leveraging the value $\sol(b)$, which yields very different values, whereas the corresponding $\Delta_{FSB}(\hat{b})$
is always $r(b) \Delta_{FSB}(b)$: 
\begin{itemize}
\item For $sol(b)<<0$, we will have 
$\Delta(\hat{b})=0<\Delta_{FSB}(\hat{b})$.
\item For $sol(b)>>0$, we will have 
$\Delta(\hat{b})=\Delta(b) > \frac{1}{2}\Delta_{FSB}(\hat{b})$ 
as $r(b)=\frac{1}{2}$.
\end{itemize}

Further, we can show that $\Utility$ is a safe overapproximation of 
$\Improve\_\max^z(a)$, which does not hold for $s_{FSB},s_{SR}$ (because of the case $sol(b)>>0$):

### 2025-04-24 ATVA 2025/proofsb.tex
final?

@@

The difference of distance to verify between SAS and GS ISO number of ReLUs is very large. However, as the ReLUs opened by GS are not as important as the ReLUs opened by SAS, the runtime of GS ISO number of ReLUs is faster than for SAS. Still, the important metric here is runtime ISO accuracy (distance to verify), which is clearly favorable for SAS, close to 2x faster. Importantly, the accuracy gain by SAS is very predictable, with improvement with each open ReLUs, while it is much more erratic for GS. This helps setting a fixed number of ReLUs. Also, SAS reaches accuracy which would be hard to match by GS in number of ReLUs to open, needing an extremely high number of open ReLUs, and thus, a uncertain high runtime.

Then, we provide Fig.~\ref{fig5}, similar to 
Fig.~\ref{fig_table3} but on a different image, namely 37.
It display similar pattern as image 85, with an even larger difference between SAS and GS (around 3x more binary variables are necessary for GS to match the accuracy of SAS).

>>>>

First, we provide Fig.~\ref{fig5}, similar to 
Fig.~\ref{fig_table3} but on a different image, namely 37.
It display similar pattern as image 85, with an even larger difference between SAS and GS (around 3x more binary variables are necessary for GS to match the accuracy of SAS).

@@

\begin{table}
	\begin{tabular}{|c||>{\centering\arraybackslash}p{6ex}|>{\centering\arraybackslash}p{6ex}|>{\centering\arraybackslash}p{6ex}|>{\centering\arraybackslash}p{6ex}|>{\centering\arraybackslash}p{6ex}|>{\centering\arraybackslash}p{6ex}|>{\centering\arraybackslash}p{6ex}|>{\centering\arraybackslash}p{6ex}|>{\centering\arraybackslash}p{6ex}||>{\centering\arraybackslash}p{6ex}||>{\centering\arraybackslash}p{6ex}|}
		\hline
		Output Node & 0 &  2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 &Avg& Var \\
		\hline
		{\sf SAS} distance & -1.47 & -4.56 & -4.48 & -3.46 & -5.6 & -3.06 & -4.17 & 0.639 & -2.33 &-3.16& \\
		Time (s) & 66.9 & 54.8 & 69.9 & 88.8 & 94.8 & 63.3 & 73.4 & 56.7 & 64.6 &70.3& 184\\
		\hline
		{\sf GS} distance & -1.38 & -4.54 & -4.45 & -3.48 & -5.7 & -2.98 & -4.08 & 0.726 & -2.28 &-3.13&\\
		Time (s) & 71.6 & 84.9 & 71.2 & 113 & 93.7 & 68.5 & 88.6 & 46.5 & 69.4 &78.6&359\\
		\hline
	\end{tabular}
		\vspace*{1ex}
		\caption{Distance to verify and runtime (lower is better) for different output nodes for image 85. Here, negative distance means the node is verified.}
		\label{SAS_GS_similarTO}
	\end{table}

Then, we provide Table \ref{SAS_GS_similarTO} to show the runtime variability of {\sf GS} and {\sf SAS}. The number of opens nodes for {\sf SAS} is 14 and the number of open nodes for {\sf GS} is 30, to account for the lack of accuracy of SAS, making the average distance to verify similar between both ($-3.16$ vs $-3.13$). The runtime variability between different output nodes is twice as high for {\sf GS} vs {\sf SAS}, because the number of irrelevant nodes chosen by {\sf GS} is inpredictable. If this number is high, then the runtime can be particularly short (node 8), with also a worse accuracy; while if it is low (e.g. 4), then the runtime is particularly long. This does not necesarily translates into better accuracy, e.g. node 2 where the runtime is much longer for {\sf GS} vs {\sf SAS}, while the accuracy is also lower. {\sf SAS} displays much less variability in time between different output nodes as most nodes are relevant.
